---
title: "Thermodurics"
author: "Caroline Motzer"
date: "2024-04-19"
output: pdf_document
---
##loading packages, setting WD/loading in data and cleaning data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(pbkrtest)
thermo <- read.csv("thermoduricdata_R.csv")
thermo$Strain<- as.factor(thermo$Strain)
thermo$Genera<- as.factor(thermo$Genera)
thermo$HeatTrt<- as.factor(thermo$HeatTrt)
thermo$IncubationTime<- as.factor(thermo$IncubationTime)
thermo$MediaType[thermo$MediaType=="SPC"]<- "SMA"
thermo$Genera[thermo$Genera=="corynebacterium"]<- "Corynebacterium"
thermo$MediaType<- as.factor(thermo$MediaType)
thermo$Comeup.Time.seconds<-period_to_seconds(ms(paste("0",thermo$Comeup.Time,sep="")))
```
##changing reference for model to microbacterium, SMA, and 48h. Also, double checking reference level is reset
```{r}
thermo<-droplevels(thermo)
thermo$MediaType<-relevel(thermo$MediaType,ref= "SMA")

thermo$Genera<-relevel(thermo$Genera, ref = "Microbacterium")

thermo$IncubationTime<-relevel(thermo$IncubationTime, ref = "48")

```
##creating log reduction column, creating "beaker" (AKA plate) column i.e linking results that came from the same plate
```{r}
thermo$log.reduction<- log10(thermo$Startingconc+1)-log10(thermo$Countcfu+1)
thermo$Beaker<- paste(thermo$Strain, thermo$HeatTrt, sep = "-")
```
##ceating model 1 AKA methodmodel and running ANOVA and checking model assumptions
```{r}
methodmodel<-lmer(log.reduction~Genera*HeatTrt*IncubationTime*MediaType+(1|Strain)+(1|Beaker), data=thermo)
summary(methodmodel)
anova(methodmodel)
hist(residuals(methodmodel))
plot(predict(methodmodel), residuals(methodmodel))
```
##creating figure 2
```{r}
thermo_filtered <- subset(thermo, IncubationTime == "48" & MediaType == "SMA")

figure2<- ggplot(thermo_filtered, aes(x = Genera, y = log.reduction)) + 
  geom_boxplot(outlier.shape = NA) +  # Remove outliers
  geom_jitter(aes(color = HeatTrt), position = position_jitter(width = 0.2), size = 2) +
  labs(y = "Reduction in Cell Count (log CFU/mL)", x = "Genera") +
  theme(axis.text.x = element_text(face = 'italic'))
figure_box2
ggsave("figure2.png", plot = figure2, width = 10, height = 6, dpi = 300)

```

##identifying what interactions are driving the significance in Model 1 AKA methodmodel
```{r}
pairs(emmeans(methodmodel,~IncubationTime))
pairs(emmeans(methodmodel,~IncubationTime|Genera, CIs=TRUE))
pairs(emmeans(methodmodel, ~MediaType))
pairs(emmeans(methodmodel,~MediaType|Genera, CIs=TRUE))
```
##Generating emip, CM I think we can delete this

```{r}
emmip(methodmodel, MediaType~IncubationTime|Genera, CIs=TRUE)
```
##creating the dataset to be used for Model 2:
##removing genera that had minimal growth (i.e., Corynebacterium, Macrococcus, Stpahylococcus) and removing 24h time point
```{r}
n=thermo%>% filter( ! (Genera%in% c("Macrococcus", "Corynebacterium", "Staphylococcus") | (IncubationTime == "24")))
n %>% view()
```


##Creating model 2 getting ANOVA and identifying which interactions are driving significance via pair wise comparisons of estimated means 
```{r}
model2<- lmer(log.reduction~Genera*HeatTrt*IncubationTime*MediaType+(1|Strain)+(1|Beaker), data=n)
anova(model2)
pairs(emmeans(model2, ~HeatTrt))
pairs(emmeans(model2, ~IncubationTime))
pairs(emmeans(model2, ~MediaType))
pairs(emmeans(model2, ~MediaType|Genera))
emmip(model2, MediaType~IncubationTime|Genera, CIs=TRUE)
```
##creating figure 4 to show differences in media and incubation time for model 2
```{r}
italic_labeller <- as_labeller(function(label) {
  paste0("italic('", label, "')")
}, default = label_parsed)


figure4 = emmip(model2, MediaType ~ IncubationTime | Genera, CIs = TRUE,
      ylab = "Reduction in Cell Count (log CFU/ml)",
      xlab = "Incubation Time (h)") +
  facet_wrap(
  ~ Genera, 
  labeller = italic_labeller)
figure4
ggsave("figure4.png", plot = figure4, dpi = 300)
```

##creating ANOVA for model 2 table to put into Word
```{r}
# Install and load necessary packages
install.packages("officer")
install.packages("flextable")
library(officer)
library(flextable)

# Example ANOVA

anova_table <- summary(model2)

# Convert the ANOVA table to a data frame
anova_mod2 <- as.data.frame(anova(model2))

anova_mod2<- anova_mod2 %>% mutate_if(is.numeric, ~ round(., 2))

# Create a flextable from the data frame
anova_flextable <- flextable(anova_mod2)


# Create a Word document
doc <- read_docx()

# Add the flextable to the document
doc <- body_add_flextable(doc, value = anova_flextable)

# Save the document
print(doc, target = "ANOVA_table1.docx")
```
##Creating Figure 5
```{r}
custom_labels <- c("A","B","C", "D")
figure5<- emmip(model2, ~HeatTrt, 
      CIs=TRUE,
      ylab = "Reduction in Cell Count (log CFU/mL)",
      xlab = expression("Heat Treatment")) +
  scale_x_discrete(labels = custom_labels)
ggsave("figure5.png", plot = figure5, dpi = 300)

```

##to make supplemental figure 1
```{r}
supplementalFigure<- ggplot(thermo,aes(x=Genera, y=log.reduction, col=Genera, shape=IncubationTime)) + 
  geom_point() + 
   ylab("Reduction in Cell Count (log CFU/mL)") +
  facet_wrap(~paste(HeatTrt,MediaType,IncubationTime)) +
   theme(axis.text.x = element_blank()) +
  guides(color = guide_legend(title.theme = element_text(face = "plain"), 
                              label.theme = element_text(face = "italic")))
supplementalFigure
ggsave("supplementalFigure.png", height = 6, width = 8, units = "in", dpi = "retina")

```
##to generate figures that relevel the 24 and 48h so 24 comes first use this model  
```{r}
z=thermo
z$IncubationTime<-relevel(z$IncubationTime, ref = "24")
z$MediaType<- relevel(z$MediaType, ref = "SMA")
z$IncubationTime
z$MediaType
zzomethodmodel<- lmer(log.reduction~Genera*HeatTrt*IncubationTime*MediaType+(1|Strain)+(1|Beaker), data=z)
anova(zzomethodmodel)
emmeans(zzomethodmodel,~IncubationTime|Genera + MediaType)
```
##Figure 3
```{r}
italic_labeller <- as_labeller(function(label) {
  paste0("italic('", label, "')")
}, default = label_parsed)

figure3<-emmip(zzomethodmodel, MediaType ~ IncubationTime | Genera, CIs = TRUE,
      ylab = "Reduction in Cell Count (log CFU/ml)",
      xlab = "Incubation Time (h)") +
  facet_wrap(
  ~ Genera, 
  labeller = italic_labeller
  )
figure3
  ggsave("figure3.png", plot = figure3, dpi = 300)
```
##to create data frame that only shows 48h counts
```{r}
thermoresults<- subset(thermo, IncubationTime == "48")
```

